---
title: 'Linear Regression Model: mtcars'
author: "Timothy Chen Allen"
date: "Tuesday, June 23, 2015"
output: html_document
---

# Multiple Regression using the `mtcars` built-in dataset
Multiple Regression extends the regression model beyond one explanatory
variable.  The purposes are still to understand relationships between
variables and make predictions.  Simple regression almost never is enough
to understand our data; mathematically, adding more explanatory variables
will make the model fit the data better.  BUT...

# Watch out for Overmodeling!
We can add more and more explanatory variables, and the model will fit the
training data better... but it may not work for other cases!  So we try to
only add variables that make sense (moon phase probably doesn't affect
wildfire occurences!) and add the least we can while still getting a good fit.

## Our Dataset: `mtcars`
`mtcars` is one of the built-in datasets in the `datasets` package of `R`.
`mtcars` data was extracted from the 1974 Motor Trend US magazine, and comprises
fuel consumption and 10 aspects of automobile design and performance for
32 automobiles (1973–74 models).

element | Description      
------- | -----------------
mpg     | Miles/(US) gallon
cyl     | Number of cylinders
disp    | Displacement (cu.in.)
hp      | Gross horsepower
drat    | Rear axle ratio
wt      | Weight (lb/1000)
qsec    | 1/4 mile time
vs      | V/Straight Engine (0 = V, 1 = straight)
am      | Transmission (0 = automatic, 1 = manual)
gear    | Number of forward gears
carb    | Number of carburetors

## Objective
Create a model of `mpg` taking into account other independent variables.

## Picking the variables
A big part of this is going to be selecting the right explanatory
variables.  This ends up being something of an art, but R facilitates it.

1) Consider that some variables in the dataset are actually dependent
variables `(mpg, hp, qsec)`, so I do not consider them.
```{r all.variable}
mt <- mtcars[c("mpg","cyl","disp","drat","wt","vs","am","gear","carb")]
# I like seeing this in mpg order to help me start understanding
(mt <- mt[order(mt$mpg,decreasing=TRUE),])
```

2) Some variables are probably highly **correlated** (e.g. `cyl` and `disp`);
we can test this using `pairs`.  Let's turn off the upper triangle,
as it's redundant.
```{r pairs}
pairs(x = mt, upper.panel=NULL, col="blue")
```

3) It appears that `cyl` and `disp`, and `disp` and `wt`, may be correlated.
Let's check it with the `cor` function
```{r cors}
cor(mt$cyl, mt$disp) # 0.902 - Just use one-- I like cyl
cor(mt$disp, mt$wt)  # 0.888 - Just use one-- I like wt
cor(mt$cyl, mt$wt)   # 0.782 - Use both
```

## Create an initial model
Let's create the model with the remaining variables, and evaluate the
model's *p*-value, Multiple R-squared, and coefficients' *p*-values.
```{r m.mpg}
m.mpg <- lm(mpg ~ cyl + drat + wt + vs + am + gear + carb, data=mtcars)
summary(m.mpg)
```
a) The *p*-value (1.489e-08) indicates the model fits the data well: good
b) The Multiple R-squared (0.853) the model explains 85.3% of the variance
in the response variable (`mpg`) based on variance in the explanatory 
variables `(cyl, drat, wt, etc)`: good
c) coefficients' *p*-values indicate they are not significant.  Not good

## Use `step` to trim down the model
Let's use `step` to strip out less useful variables.  `step` is pretty
dumb; it just tries every possible model and chooses the one with the
lowest AIC (*Akaike Information Criteria*).  Don't just let `step` create
your model for you; you need to understand possible relationships or the
model becomes meaningless. `step` also produces a lot of output
if you don't specify `trace=0`.
```{r}
m.mpg.step <- step(object = m.mpg, direction = "both", trace=0)
summary(m.mpg.step)
```

although `carb` appears, its coefficient is not significant.
build a model without it:
```{r mpg2}
m.mpg2 <- lm(mpg ~ cyl + wt, data=mtcars)
summary(m.mpg2)
```

## Tweak the model: interaction effects
I want to try out a single interaction effect; this means that I 
believe there is a possibility that the combination of less cylinders
and a manual transmission may improve mileage more than those
factors by themselves.

This points out an unfortunate fact: it is easier to build a model
if you have some understanding of the subject matter.
```{r m.mpg3}
m.mpg3 <- lm(mpg ~ cyl * am + wt, data=mtcars)
summary(m.mpg3)
par(mfrow=c(2,2))
plot(m.mpg3, col="blue")
```

## Compare models with the AIC
The **Akaike Information Criteria** is a way of comparing models.
It rewards better fit, but penalizes additional explanatory
variables (thus avoiding *overmodeling*):
```{r AIC}
AIC(m.mpg)
AIC(m.mpg.step)
AIC(m.mpg2)
AIC(m.mpg3)
```

## Pick a model
I like `m.mpg3 (mpg ~ cyl * am + wt)`; it makes some logical sense:
a light car with a smaller engine and manual transmission
would seem to have better gas mileage.
All coefficients are significant, and the AIC is the lowest of
the four.
```{r pick.one}
m.mpg3
```

## Use the model to predict
The 2015 Kia Soul with a manual transmission gets 26 mpg.
The model predicts 25.94; pretty good.
```{r predict}
kia.soul <- data.frame(mpg=26, cyl=4, am=1, wt=2.7)
row.names(kia.soul) <- "Kia Soul 2015"
predict(m.mpg3, newdata=kia.soul)
```


