---
title: "Simple Regression Model"
author: "Timothy Chen Allen"
date: "Wednesday, June 24, 2015"
output: html_document
---

# Simple Data Model
This is a "pretend" data model based on a known relationship: F to C
We'll randomize this relationship a little and show how we model it.  

## Purpose of Modeling
Remember: the purposes of modeling are:
Based on historical data...
1) understand relationships between variables
2) allow prediction 

## Our Fake Data
1) Create a model using a known formula (Fahrenheit to Celsius),
but include a random effect (maybe we're looking at two different
thermometers a block apart).
```{r fake.data}
N <- 4000
f <- rnorm(n=N, mean=45, sd=20)
c <- (f-32)*(5/9) + rnorm(n=N, mean=0, sd=10)
```

## Let's see some of the data
```{r inspect.data}
head(data.frame(f,c),10)
```

## Create a model predicting Celsius given Fahrenheit
We expect the model to look a lot like the original
formula.  Draw a line with the model's prediction values.
```{r create.model}
plot(x=f, y=c, main=paste("Simple Linear Regression:", N), pch=21,
     bg=heat.colors(139)[round(f-min(f))+1])
# This creates the model (parentheses around the command displays the output)
(m <- lm(c ~ f))
abline(m, col="purple", lty="dashed", lwd=3)

# Draw another line with the predicted values of the perfect formula
perfect.f <- 0:80
perfect.c <- (perfect.f-32)*(5/9)
abline(lm(perfect.c ~ perfect.f), col="green", lty="dashed", lwd=3)
```

## Diagnose the model with `summary` and `plot`
```{r diagnose}
summary(m)
par(mfrow=c(2,2))
plot(m, col="blue")
```

## Use the model to predict
We expect 95F to predict about 35C; the model predicts 34.6, so pretty good.
```{r predict}
test.case <- data.frame(f=95) 
predict(m, newdata=test.case)
```