---
title: "Simple Regresion - Pressure vs Speed"
author: "Timothy Chen Allen"
date: "Tuesday, August 04, 2015"
output: html_document
---

# Simple Data Model
This is a data model based on a known relationship: windspeed vs
barometric pressure within a hurricane.  These data combined
FEMA declaration data with NOAA weather data.

## Purpose of Modeling
Remember: the purposes of modeling are:
Based on historical data...
1) understand relationships between variables
2) allow prediction 

## Get the data
These data combine FEMA declaration data with NOAA storm data.
```{r get.data}
setInternet2(use=TRUE)
URL <- "https://raw.githubusercontent.com/TimothyChenAllen/ten_tasks/master/speed_vs_pressure.csv"
download.file(URL, destfile = "speed_vs_pressure.csv", mode="wb")
storms <- read.csv("speed_vs_pressure.csv")
```

## Let's see some of the data
`speed` in knots, `pressure` in mbars
```{r inspect.data}
head(storms,10)
```

## Create a model predicting Celsius given Fahrenheit
We know that lower pressure is related to higher windspeed.
Plot the data, create the model, and compare predicted values to actuals
Draw a line with the model's prediction values.
```{r create.model}
par(mfrow=c(1,1))
plot(speed ~ pressure, data=storms, pch=cat+20, xlab="Pressure (mbars)", 
      ylab="Windspeed (knots)", bg=rainbow(7)[cat+2], col="gray50",
      main="Hurricane Windspeed vs Pressure")

# This creates the model
m.storms <- lm(speed ~ pressure, data=storms)

# Draw the model's predicted values against the observations
abline(m.storms, col="purple", lty="dashed", lwd=3)
```

## Diagnose the model with `summary` and `plot`
```{r diagnose}
summary(m.storms)
par(mfrow=c(2,2))
plot(m.storms, col="blue")
```

## Use the model to predict
Katrina was 902 mbars and 175 mph
```{r predict}
test.case <- data.frame(pressure=902) 
predict(m.storms, newdata=test.case)
```

# Make a more advanced plot using ggplot2
```{r ggplot}
require(ggplot2)
par(mfrow=c(1,1))
ggplot(storms, aes(x=pressure, y=speed, fill=rainbow(7)[cat+2])) +
    geom_point(shape=storms$cat+20) +
    scale_colour_hue(l=10) +
    geom_smooth(aes(x=pressure, y=speed, colour="grey", fill="black", alpha=.1), data=storms, method=lm) +
    theme(legend.position="none") + 
    ggtitle("Storm Windspeed vs Pressure") + 
    xlab("Pressure (mbar)") +
    ylab("Windspeed (knots)") +
    theme(panel.background = element_blank()) +
    geom_text(aes(x=890,y=50, hjust=0,label="Speed ~ 1241.5 - 1.194 x Pressure", fontface=3)) +
    geom_text(aes(x=965,y=175, hjust=0,label="p-value = 2.2e-16", fontface=3)) +
    geom_text(aes(x=965,y=155, hjust=0,label="R^2 = 0.9271", fontface=3))
```